
---
title: "MLBuoy R Package Vignette"
output:
  prettydoc::html_pretty:
    theme: HPSTR
vignette: >
  %\VignetteIndexEntry{MLBuoy-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE, 
  message = FALSE, 
  warning = FALSE
)
```

# **Introduction**

Grand Valley State University's Robert B. Annis Water Resources Institute (AWRI) established a buoy-based observatory in Muskegon Lake in 2010. The Muskegon Lake Buoy is an important tool for monitoring the conditions of the lake and providing valuable information about the environment. It provides current conditions and historic data of Muskegon Lake in Muskegon County, Michigan going back to 2011, enabling researchers and stakeholders to gain a comprehensive understanding of the lake's ecosystem over time. 

In order to facilitate ease of extraction and visualization of the data, we developed an R Package called MLBuoy that utilizes an Application Programming Interface (API) to provide users with an efficient way to access and work with the data from the Muskegon Lake Buoy. The R Package includes various functions and tools that enable users to easily extract, manipulate, and visualize the data, allowing for in-depth analysis and insights. By utilizing an API, the MLBuoy package makes it simple for users to access and use the data in their own applications, software, and workflows.

The first function, fetch_buoy_data, obtains data on current and historical conditions of Muskegon Lake in Muskegon County, Michigan from the Muskegon Lake Buoy API. The second function, plot_buoy_data, offers a simple method to visualize data obtained from Muskegon Lake Buoy.


# **Function 1: Fetch Muskegon Lake Buoy Data**

Fetch Muskegon Lake Buoy Data is used to retrieve data from the Muskegon Lake Buoy. It allows the user to obtain valuable information about the current conditions and historical data of the lake, all through the convenience of an API. The function provides comprehensive understanding of the lake's conditions over time as it grants access to a wealth of data dating back to 2011.

## **Parameters**

The function takes the following parameters:

```{r, echo=FALSE}
library(knitr)
kable(
  data.frame(
    Variable = c("x", "y", "start_date",  "end_date", "time", "calculation"), 
    Description = c("A single x value", "At least one y value is required", "A specific date in time that marks the beginning of a particular period","A specific date in time that marks the end of a particular period", "Time values must be in the 24-hour format H:MM or HH:MM.", "The mathematical calculation performed on the range points within an x variable's concentration"),
    Supported_Values = c("If no value is specified, date will be used as the default", "Multiple y values should be separated by a comma", "Date values must be in the format MM/DD/YYYY or M/D/YY. The default value is 10/01/2022", "Values must be in the format MM/DD/YYYY or M/D/YY. The default date is 10/31/2022", "The values must be divisible by 15 minutes (0:00, 0:15, 0:30, and 0:45). If no value is provided then all times are considered.", "Average, Minimum and Maximum")), 
  format = "pandoc"
)
```


Note: The arguments must use supported values. More information can be found at Grand Valley State University Muskegon Lake Buoy API website: https://www.gvsu.edu/wri/buoy/data-api.htm

## **Function Output**

The Fetch Muskegon Lake Buoy Data function returns a data frame based on the provided parameters.

## **Required Libraries**

Before using the Fetch Muskegon Lake Buoy Data function, you must import tidyverse library.

## **Example**

Here is an example of how to use the Fetch Muskegon Lake Buoy Data function:

```{r, error=FALSE}
# Installation the package: devtools::install_github("https://github.com/BNgigi/MLBuoy")
# Load the libraries: tidyverse and MLBuoy
library(MLBuoy)
library(tidyverse)

buoyData <- fetch_buoy_data(x="weekday",y="atmp1,tp001,tp002", start_date="7/7/11", end_date="7/28/11")

knitr::kable(head(buoyData, n = 12))
```

In this example, we are retrieving data for the atmp1, tp001, and tp002 as per day of week for the period between July 7, 2011 and July 28, 2011.

# **Function 2: Plot Muskegon Lake Buoy Data**

Plot Muskegon Lake Buoy Data is designed to retrieve and extracts data from the Muskegon Lake Buoy through the use of an Application Programming Interface (API). The raw data is then transformed into an easy-to-read and visually appealing graph, facilitating comprehension and analysis.

## **Parameters**

The function takes the following parameters:

```{r, echo=FALSE}
library(knitr)
kable(
  data.frame(
    Variable = c("x", "y", "start_date", "end_date","graph_type"), 
    Description = c("The values to be plotted on the x-axis of the graph. If no value is specified, date will be used as the default.", "The values to be used on the y-axis of the graph. Multiple y values should be separated by a comma.","A specific date in time that marks the beginning of a particular period. Date values must be in the format MM/DD/YYYY or M/D/YY. The default value is 10/01/2022.", "A specific date in time that marks the end of a particular period. Values must be in the format MM/DD/YYYY or M/D/YY. The default date is 10/31/2022.", "The type of chart to be plotted. The supported types are scatter, line, bar, and boxplot.")), 
  format = "pandoc"
)
```


Note: The arguments for x and y must be supported values. For more information, please visit the Grand Valley State University Muskegon Lake Buoy API website: https://www.gvsu.edu/wri/buoy/data-api.htm

## **Function Output**

The Plot Muskegon Lake Buoy Data function returns a graph of x values by y values.

## **Importing Required Libraries**

Before using the Plot Muskegon Lake Buoy Data function, you must import tidyverse library.

## **Function Usage**

The following is an example of how to use the Plot Muskegon Lake Buoy Data function:

```{r, error=TRUE}
library(tidyverse)
library(MLBuoy)

plot_buoy_data (x="weekday", y="rh1", start_date="7/7/11", end_date="7/28/11", graph_type="scatter")

plot_buoy_data (x="weekday", y="rh1", start_date="7/7/11", end_date="7/28/11", graph_type="line")

plot_buoy_data (x="weekday", y="rh1", start_date="7/7/11", end_date="7/28/11", graph_type="bar")

plot_buoy_data (x="weekday", y="rh1", start_date="7/7/11", end_date="7/28/11", graph_type="boxplot")
```

# **Package Development**

The MLBuoy R Package was developed by Beatrice Ngigi and Andrew DiLernia.


